{\fontfamily{ptm}\selectfont A Parameter Free Accelerated Proximal Gradient Method Without Restarting }

### 2025-01-03 Notebook LaTEX/Proposal/main.tex
A lot of changes

\section{Preliminaries}
    \todoinline{
        Clarify: Notations,
        Organizations. 
    }
    This section contains the basics of contents from convex optimization, and variational analysis. 
    \par
    Notations. 
    \begin{enumerate}
        \item $\overline \RR := \RR \cup \{\infty, -\infty\}$
    \end{enumerate}
    \subsection{Fundamentals in non-convex analysis}
        \todoinline{\noindent
            We are in $\RR^n$, and the weakest assumption we are making for the objective function is Local Lipschitz continuity. 
            Definitions: 
            \begin{enumerate}
                \item[$\blacksquare$] Local Lipschitz continuity. 
                \item[$\square$]Regular subgradient. Remember to cite. 
                \item[$\square$] Limiting subgradient. Remember to cite. 
                \item[$\blacksquare$] Weakly convex function. 
                \item[$\blacksquare$]The Bregman Divergence of function. 
            \end{enumerate}
            Take Limiting, Regular subgradient definitions from Cui, Pong's book, Definition 4.3.1. 
        }
        Let the ambient space be $\RR^n$ equipped with inner product and 2-norm. 
        Let $O$ be an open subset of $\RR^n$, the weakest assumption we are making for the objective function $F: O \subseteq \RR^n \rightarrow \RR$ for optimization problem is Local Lipschitz Continuity. 
        The assumption of local Lipschitz continuity is weak enough to describe most problems in applications, and strong enough to avoid most pathologies in analysis.

@@

            Introduce 
                \item[$\blacksquare$] Convexity, 
                \item[$\blacksquare$] convex subgradient, 
                \item[$\blacksquare$] Lipschitz smoothness. 
                \item [$\blacksquare$]Strong convexity of a function. 
                \item [$\blacksquare$]The proximal gradient operator. 
                \item [$\blacksquare$]The proximal mapping operator. 
            Lemmas: 
                \item [$\blacksquare$]Quadratic growth conditions of a strongly convex function. 

@@

        \subsubsection{Smooth, nonsmooth additive composite}\label{sssec:additive-composite}
            \todoinline{\noindent
                Introduce notations for the proximal gradient model function. 
                Lemmas: 
                \begin{enumerate}
                    \item[$\blacksquare$] Proximal gradient envelope. 
                    \item[$\blacksquare$] A property of gradient mapping. 
                \end{enumerate}
                Theorems: 
                \begin{enumerate}
                    \item[$\blacksquare$] The proximal gradient inequality. 
                \end{enumerate}
            }

            In this section, we zoom in further. 
            Suppose that $F:= f + g$ where $f:\RR^n \rightarrow \RR$ is convex, $L$ Lipschitz smooth and $\mu \ge 0$ strongly convex and $g: \RR\rightarrow \overline \RR$ is convex. 
            To make the discussion simpler, fix any $\beta \ge 0$ we define the following model functions as a $\RR^n \times \RR^n \rightarrow \overline \RR$: 
            \begin{align*}
                \widetilde{\mathcal M}^{\beta^{-1}}
                (x; y)
                &:= 
                g(x) + f(y) + \langle \nabla f(y), x - y\rangle
                + \frac{\beta}{2}\Vert x - y\Vert^2,
                \\
                \mathcal M^{\beta^{-1}}(x; y) 
                &:= F(x) + \frac{\beta}{2}\Vert x - y\Vert^2.
            \end{align*}
            Under convexity assumption in this section, both $\widetilde {\mathcal M} (\cdot; y),  {\mathcal M}(\cdot;y )$ is at least $\beta \ge 0$ strongly convex. 
            \begin{definition}[Proximal gradient operator]\label{def:proximal-gradient-operator}
                Take $F := f + g$ where $g: \RR^n \rightarrow \overline \RR$ as defined in this section. 
                Define the proximal gradient operator $T_L$ on all $y \in \RR^n$: 
                \begin{align*}
                    T_L y := \argmin_{x \in \RR^n} \left\lbrace
                        g(x) + f(y) + \langle \nabla f(y), x - y\rangle 
                        + \frac{L}{2}\Vert x- y \Vert^2
                    \right\rbrace. 
                \end{align*}
            \end{definition}
            \begin{remark}
                Under the assumption of this section, the mapping $T_L$ is a single-valued mapping, it has domain on the entire $\RR^n$, and it's a $3/2$ averaged operator. 
            \end{remark}
            \begin{definition}[Gradient mapping operator]\label{def:gradient-mapping-operator}
                Take $F := f + g$ as defined in this section. 
                Define the gradient mapping operator $\mathcal G_L$ on all $y \in \RR^n$: 
                \begin{align*}
                    \mathcal G_L y:= L(y - T_L y). 
                \end{align*}
            \end{definition}

@@

\section{Catalyst accelerations and future works}\label{sec:catalyst}
    \todoinline{\noindent
        Literatures review of the topics in Catalyst acceleration method. 
        Here is a list of topics: 
        \begin{enumerate}
            \item The original accelerated PPM. 
            \item The Catalyst with weakly convex objectives. 
        \end{enumerate}
        After the literature reviews of the core literatures, move on and state new research directions and open problems. 
        There are several directions for open problem: 
        \begin{enumerate}
            \item APPM method for monotone operators instead of just subgradient, whether the same framework exists in a greater context. 
            \item Accelerated Proximal Bregman Method. 
            \item Removing smoothness assumption in Catalyst acceleration framework. 
        \end{enumerate}  
    }
    \todoinline{\noindent
        A list of relevant literatures: 
        \begin{enumerate}
            \item Güler's 1992 paper on Accelerated Proximal Point method. 
            \item Lin's, and Payquette's three trilogy paper on Catalyst acceleration for convex, non-convex Variance reduced algorithm. 
        \end{enumerate}
    }
    In this section, we assume that $F: \RR^n \rightarrow \overline \RR$ is a convex function. 
    This section reports major results from literatures concerning the method of Catalyst Acceleration. 
    \begin{assumption}
        Given any $\beta > 0$ and $y \in \RR^n$, and $F: \RR^n\rightarrow \overline \RR$ is $\mu \ge 0$ strongly convex and closed. 
        Define the model function for all $y\in \RR^n$ to be 
        \begin{align*}
            \mathcal M^{\beta^{-1}}(\cdot; y) &:= 
            F(x) + \frac{\beta}{2}\Vert x - y\Vert^2.
        \end{align*}
        We define the Moreau Envelope at $y \in \RR^n$ to be $\mathcal M^*_{\beta^{-1}}(y) = \min_{x\in \RR^n} \mathcal M^{\beta^{-1}}(x; y)$. 
        We denote $\mathcal J_{\beta^{-1}}$ to be the resolvent operator for subgradient of $F$, and $\mathcal J_{\beta^{-1}}^\epsilon$ for any $\epsilon \ge 0$ to be the inexact resolvent operator. 
        \begin{align*}
            \mathcal J_{\beta^{-1}}^\epsilon y &:= 
            \left\lbrace
                x \in \RR \left | 
                    \mathcal M^{\beta^{-1}}(x; y) - 
                    \mathcal M_{\beta^{-1}}^*(y) \le \epsilon
                \right.
            \right\rbrace. 
        \end{align*}
        Setting $\epsilon = 0$, we have the exact definition of the exact resolvent given as $\mathcal J_{\beta^{-1}}y = \mathcal J^0_{\beta}y$. 
        Finally, suppose that $f: \RR^n \rightarrow \RR$ is a $L$-Lipschitz smooth and $\mu \ge 0$ strongly convex function, and $g: \RR^n \rightarrow \overline \RR$ is a closed convex function, we define the linearized model function: 
        \begin{align*}
            \widetilde{\mathcal M}^{\beta^{-1}}(x; y) := 
            g(x) + f(y) + \langle \nabla f(y), x - y\rangle + \frac{\beta}{2}\Vert x - y\Vert^2. 
        \end{align*}
    \end{assumption}

    Inspired by accelerated proximal point method from Güler \cite{guler_new_1992}, and inexact proximal point method of Rockafellar 1976 \cite{rockafellar_monotone_1976}, Lin \cite{lin_universal_2015} proposed a generic method taking inspirations from the convergence claims of Accelerated proximal point method to accelerated the convergence rate of first order variance reduced incremental method. 
    The class of variance reduced method is vast, but to use the most relevant feature of this class of first order method is that they are stochastic method that is not slower than full gradient descent in complexity. 
    \todo{
        $\blacksquare$ Add the bib for Mark Schmidt introductory paper RV methods.  
        \\ $\square$ Cite it. 
    }{\hl{See Gower's guide}} for more information on variance reduced methods in machine learning. 
    \par
    The parts coming will introduce the algorithms, explain key innovations behind the algorithm, and key variants of the algorithm. 
    The section at the end will give future works and extensions of the Catalyst framework.

    \begin{definition}[Lin's Universal Catalyst Acceleration]\label{def:lin-catalyst}\; \\
        Let the initial estimate be $x_0 \in \RR^n$, fix parameters $\kappa > 0$ and $\alpha_0 \in (0, 1]$. 
        Let $(\epsilon_k)_{k \ge 0}$ be an error sequence chosen for the evaluation for inexact proximal point method. 
        Initialize $x_0 = y_0$. Then the algorithm generates $(x_k, y_k)_{k\ge 0}$ for all $k \ge 1$ such that: 
        \begin{align*}
            & \text{find } x_k \in \mathcal J_{\kappa^{-1}}^{\epsilon_k} y_{k - 1}, 
            \\
            & \text{find } \alpha_k \in (0, 1) \text{ such that } \alpha_k^2 = (1 - \alpha_k)\alpha_{k - 1}^2 + (\mu/(\mu + \kappa))\alpha_k,
            \\
            & 
            y_{k} = x_k + \frac{\alpha_{k - 1}(1 - \alpha_{k - 1})}{\alpha_{k - 1}^2 + \alpha_k}(x_k - x_{k - 1}). 
        \end{align*}
    \end{definition}
    \begin{remark}
        The above algorithm is Algorithm 1 from Lin's first paper on Catalyst Acceleration \cite{lin_universal_2015}. 
        The explicit formula for $\alpha_k$ is the larger root of solving the quadratic equation given by: 
        \begin{align*}
            \alpha_k &= 
            \frac{1}{2}\left(
                - \alpha_{k - 1}^2 - q + \sqrt{(q + \alpha_{k - 1})^2 + 4 \alpha_{k - 1}}
            \right), 
        \end{align*}
        where $q = \mu/(k + \mu)$. 
        Lin suggests different choices for the parameter $\kappa > 0$ depending on the algorithm chosen to evaluate the subroutine for $\mathcal J_{\kappa^{-1}}^{\epsilon_k}y_{k - 1}$. 
        The choice of $\epsilon_k$ depends on the estimated optimality gap $F(x_0) - F^*$ where $F^*$ is the minimum of $F$ and whether $\mu > 0$ or $\mu = 0$. 
    \end{remark}
    Much of the heavy lifting of constructing the algorithms were done through the method of estimating sequence introduced back in Definition \ref{def:nes-est-seq}. 
    The convergence of the algorithm made use of an inexact version of the proximal gradient inequality (similar to Theorem \ref{thm:prox-grad-ineq}) stated as Lemma A.7 in \cite{lin_universal_2015}. 
    This lemma is instrumental for deriving an inexact variant of the estimating sequence $\phi_k^* \ge F(x_k) + \xi_k$. 
    The sequence $\epsilon_k$ is compacted and bounded under an upper bound parameterized via the sequence $(\epsilon_k)_{k \ge 0}$. 
    The convergence proof from Lin was inspired by 
    \todo{\noindent
        $\blacksquare$ Add bib, the paper is 
        \\
        $\square$ Cite. 
    }{\hl{Schmidt's Inexact Proximal Gradient method. }} 
    \par
    The technique of Estimating sequence results in depressingly long proof making it unsuitable for exposition here, however every significant piece of innovation is covered in details in our most recent Fall Winter 2024 MATH 590 report. 
    \par 
    The following parts will complement the report and describe ideas on Lin's idea for choosing,termination criteria of the inner loop to evaluate $\mathcal J_k^{\epsilon_k} y_k$ in his second paper on Catalyst Acceleration \cite{lin_catalyst_2018}.

>>>>

\section{Preliminaries}
    \todoinline{
        Clarify: Notations,
        Organizations. 
    }
    This section contains the basics of contents from convex optimization, and variational analysis. 
    \par
    Notations. 
    \begin{enumerate}
        \item $\overline \RR := \RR \cup \{\infty, -\infty\}$
    \end{enumerate}
    \subsection{Fundamentals in non-convex analysis}
        Let the ambient space be $\RR^n$ equipped with inner product and 2-norm. 
        Let $O$ be an open subset of $\RR^n$, the weakest assumption we are making for the objective function $F: O \subseteq \RR^n \rightarrow \RR$ for optimization problem is Local Lipschitz Continuity. 
        The assumption of local Lipschitz continuity is weak enough to describe most problems in applications, and strong enough to avoid most pathologies in analysis.

@@

@@

        \subsubsection{Smooth, nonsmooth additive composite}\label{sssec:additive-composite}
            In this section, we zoom in further. 
            Suppose that $F:= f + g$ where $f:\RR^n \rightarrow \RR$ is convex, $L$ Lipschitz smooth and $\mu \ge 0$ strongly convex and $g: \RR\rightarrow \overline \RR$ is convex. 
            To make the discussion simpler, fix any $\beta \ge 0$ we define the following model functions as a $\RR^n \times \RR^n \rightarrow \overline \RR$: 
            \begin{align*}
                \widetilde{\mathcal M}^{\beta^{-1}}
                (x; y)
                &:= 
                g(x) + f(y) + \langle \nabla f(y), x - y\rangle
                + \frac{\beta}{2}\Vert x - y\Vert^2,
                \\
                \mathcal M^{\beta^{-1}}(x; y) 
                &:= F(x) + \frac{\beta}{2}\Vert x - y\Vert^2.
            \end{align*}
            Under convexity assumption in this section, both $\widetilde {\mathcal M} (\cdot; y),  {\mathcal M}(\cdot;y )$ is at least $\beta \ge 0$ strongly convex. 
            \begin{definition}[Proximal gradient operator]\label{def:proximal-gradient-operator}
                Take $F := f + g$ where $g: \RR^n \rightarrow \overline \RR$ as defined in this section. 
                Define the proximal gradient operator $T_L$ on all $y \in \RR^n$: 
                \begin{align*}
                    T_L y := \argmin_{x \in \RR^n} \left\lbrace
                        g(x) + f(y) + \langle \nabla f(y), x - y\rangle 
                        + \frac{L}{2}\Vert x- y \Vert^2
                    \right\rbrace. 
                \end{align*}
            \end{definition}
            \begin{remark}
                Under the assumption of this section, the mapping $T_L$ is a single-valued mapping, it has domain on the entire $\RR^n$, and it's a $3/2$ averaged operator. 
            \end{remark}
            \begin{definition}[Gradient mapping operator]\label{def:gradient-mapping-operator}
                Take $F := f + g$ as defined in this section. 
                Define the gradient mapping operator $\mathcal G_L$ on all $y \in \RR^n$: 
                \begin{align*}
                    \mathcal G_L y:= L(y - T_L y). 
                \end{align*}
            \end{definition}

@@

\section{Catalyst accelerations and future works}\label{sec:catalyst}
    \todoinline{\noindent
        Literatures review of the topics in Catalyst acceleration method. 
        Here is a list of topics: 
        \begin{enumerate}
            \item The original accelerated PPM. 
            \item The Catalyst with weakly convex objectives. 
        \end{enumerate}
        After the literature reviews of the core literatures, move on and state new research directions and open problems. 
        There are several directions for open problem: 
        \begin{enumerate}
            \item APPM method for monotone operators instead of just subgradient, whether the same framework exists in a greater context. 
            \item Accelerated Proximal Bregman Method. 
            \item Removing smoothness assumption in Catalyst acceleration framework. 
        \end{enumerate}  
    }
    \todoinline{\noindent
        A list of relevant literatures: 
        \begin{enumerate}
            \item Güler's 1992 paper on Accelerated Proximal Point method. 
            \item Lin's, and Payquette's three trilogy paper on Catalyst acceleration for convex, non-convex Variance reduced algorithm. 
        \end{enumerate}
    }
    In this section, we assume that $F: \RR^n \rightarrow \overline \RR$ is a convex function. 
    This section reports major results from literatures concerning the method of Catalyst Acceleration. 
    \begin{assumption}
        Given any $\beta > 0$ and $y \in \RR^n$, and $F: \RR^n\rightarrow \overline \RR$ is $\mu \ge 0$ strongly convex and closed. 
        Define the model function for all $y\in \RR^n$ to be 
        \begin{align*}
            \mathcal M^{\beta^{-1}}(\cdot; y) &:= 
            F(x) + \frac{\beta}{2}\Vert x - y\Vert^2.
        \end{align*}
        We define the Moreau Envelope at $y \in \RR^n$ to be $\mathcal M^*_{\beta^{-1}}(y) = \min_{x\in \RR^n} \mathcal M^{\beta^{-1}}(x; y)$. 
        We denote $\mathcal J_{\beta^{-1}}$ to be the resolvent operator for subgradient of $F$, and $\mathcal J_{\beta^{-1}}^\epsilon$ for any $\epsilon \ge 0$ to be the inexact resolvent operator. 
        \begin{align*}
            \mathcal J_{\beta^{-1}}^\epsilon y &:= 
            \left\lbrace
                x \in \RR \left | 
                    \mathcal M^{\beta^{-1}}(x; y) - 
                    \mathcal M_{\beta^{-1}}^*(y) \le \epsilon
                \right.
            \right\rbrace. 
        \end{align*}
        Setting $\epsilon = 0$, we have the exact definition of the exact resolvent given as $\mathcal J_{\beta^{-1}}y = \mathcal J^0_{\beta}y$. 
    \end{assumption}

    Inspired by accelerated proximal point method from Güler \cite{guler_new_1992}, and inexact proximal point method of Rockafellar 1976 \cite{rockafellar_monotone_1976}, Lin \cite{lin_universal_2015} proposed a generic method taking inspirations from the convergence claims of Accelerated proximal point method to accelerated the convergence rate of first order variance reduced incremental method. 
    The class of variance reduced method is vast, but to use the most relevant feature of this class of first order method is that they are stochastic method that is not slower than full gradient descent in complexity. 
    \todo{
        $\blacksquare$ Add the bib for Mark Schmidt introductory paper RV methods.  
        \\ $\square$ Cite it. 
    }{\hl{See Gower's guide}} for more information on variance reduced methods in machine learning. 
    In brief, a variance reduce method (VRM) is a type of stochastic gradient methods that stablizes the current estimate using information of the gradient at previous iterates. 
    In each iteration, only the gradient of a few samples are accessed but attain overall better complexity because the variance of the estimated gradients are reduced. 
    \todoinline{
        We can then list some citations and references to commonly known variance reduced method here. 
    }
    \par
    The parts coming will introduce the algorithms, explain key innovations behind the algorithm, and key variants of the algorithm. 
    The section at the end will give future works and extensions of the Catalyst framework.

    \begin{definition}[Lin's Universal Catalyst Acceleration]\label{def:lin-catalyst}\; \\
        Let the initial estimate be $x_0 \in \RR^n$, fix parameters $\kappa > 0$ and $\alpha_0 \in (0, 1]$. 
        Let $(\epsilon_k)_{k \ge 0}$ be an error sequence chosen for the evaluation for inexact proximal point method. 
        Initialize $x_0 = y_0$. Then the algorithm generates $(x_k, y_k)_{k\ge 0}$ for all $k \ge 1$ such that: 
            & \text{find } x_k \in \mathcal J_{\kappa^{-1}}^{\epsilon_k} y_{k - 1}, 
            & \text{find } \alpha_k \in (0, 1) \text{ such that } \alpha_k^2 = (1 - \alpha_k)\alpha_{k - 1}^2 + (\mu/(\mu + \kappa))\alpha_k,
            y_{k} = x_k + \frac{\alpha_{k - 1}(1 - \alpha_{k - 1})}{\alpha_{k - 1}^2 + \alpha_k}(x_k - x_{k - 1}). 
        The above algorithm is Algorithm 1 from Lin's first paper on Catalyst Acceleration \cite{lin_universal_2015}. 
        The explicit formula for $\alpha_k$ is the larger root of solving the quadratic equation given by: 
            \alpha_k &= 
            \frac{1}{2}\left(
                - \alpha_{k - 1}^2 - q + \sqrt{(q + \alpha_{k - 1})^2 + 4 \alpha_{k - 1}}
        where $q = \mu/(k + \mu)$. 
        Lin suggests different choices for the parameter $\kappa > 0$ depending on the algorithm chosen to evaluate the subroutine for $\mathcal J_{\kappa^{-1}}^{\epsilon_k}y_{k - 1}$. 
        The choice of $\epsilon_k$ depends on the estimated optimality gap $F(x_0) - F^*$ where $F^*$ is the minimum of $F$ and whether $\mu > 0$ or $\mu = 0$. 
    Much of the heavy lifting of constructing the algorithms were done through the method of estimating sequence introduced back in Definition \ref{def:nes-est-seq}. 
    The convergence of the algorithm made use of an inexact version of the proximal gradient inequality (similar to Theorem \ref{thm:prox-grad-ineq}) stated as Lemma A.7 in \cite{lin_universal_2015}. 
    This lemma is instrumental for deriving an inexact variant of the estimating sequence $\phi_k^* \ge F(x_k) + \xi_k$. 
    The sequence $\epsilon_k$ is compacted and bounded under an upper bound parameterized via the sequence $(\epsilon_k)_{k \ge 0}$. 
    The convergence proof from Lin was inspired by 
    \todo{\noindent
        $\blacksquare$ Add bib, the paper is 
        $\square$ Cite. 
    }{\hl{Schmidt's Inexact Proximal Gradient method. }} 
    The technique of Estimating sequence results in depressingly long proof making it unsuitable for exposition here, however every significant piece of innovation is covered in details in our most recent Fall Winter 2024 MATH 590 report. 
    The following parts will complement the report and describe ideas on Lin's idea for choosing, termination criteria of the inner loop to evaluate $\mathcal J_k^{\epsilon_k} y_k$ in his second paper on Catalyst Acceleration \cite{lin_catalyst_2018}. 
    To elucidate the matters consider the model function $\mathcal M^{\kappa^{-1}}(x; y)$ is $\mu + \kappa$ strongly convex, and therefore it admits the Polyak error bound condition: 
        (\forall x\in \RR^n) \quad 
        \mathcal M^{\kappa^{-1}}(x; y) -
        \mathcal M_{\kappa^{-1}}^*(y) 
        &\le (1 + \mu)\dist
            \left(\mathbf 0, \partial \mathcal M^{\kappa^{-1}}(x; y)
    By choosing $x$ such that $\dist(\mathbf 0, \partial \mathcal M^{\kappa^{-1}}(x; y)) \le \epsilon$, it ensures $x \in \mathcal J^\epsilon_{\kappa^{-1}}(y)$. 
    Unfortunately in practice, this is not used because a full gradient evaluation on the model function $\mathcal M^{\kappa}(\cdot; y)$ is costly (compare to the small amount required for variance reduced incremental method), so Lin suggested alternatives of Inner Loop Termination Criteria to make Catalyst acceleration competitive in practice. 
        Elucidate things by considering: 
            \item What is the suggested error sequence $\epsilon_k$ for different type of objective function? 
            \item What are some conditions for achieving inexact evaluation up to the given accuracy $\epsilon_k$? 
            \item What are the convergence claims for these different types of termination criteria. 
    \subsection{Error sequence and convergence claims}
    \subsection{Inner loop termination criteria}    

### 2025-01-03 Notebook LaTEX/R-WAPG/paper3.tex
A lot of changes

            This section present results of numerical experiment for solving the LASSO problem proposed by 

@@

                $\square$ Add reference ``Regression shrinkage and selection via the Lasso''
                \\ $\square$ Cite it. 
            }{\hl{Tibshirani}}. 

>>>>

            This section present results of numerical experiment for solving the (least absolute shrinkage and
            selection operator) LASSO problem proposed by 

@@

                $\blacksquare$ Add reference ``Regression shrinkage and selection via the Lasso''
                \\ $\blacksquare$ Cite it. 
            }{\hl{Tibshirani}} \cite{tibshirani_regression_1996}. 
